Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-09 paper/paper.tex
figure fixed

\documentclass[9pt]{sigplanconf}
\usepackage{etex}
\usepackage{float}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage{boxedminipage}
\usepackage[T1]{fontenc}
%\usepackage[pdf]{pstricks}
\usepackage{epsfig}
\usepackage{multirow}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{schemepgm}
\usepackage{graphicx}
\usepackage{graphpap}
\usepackage{tabularx}
\usepackage [plain,noend,noline,boxed]{algorithm2e}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{pstricks}
\usepackage{pst-text}
\usepackage{pst-node}
\usepackage{pst-tree}
\usepackage{pst-rel-points}
\usepackage{bcprules}
\usepackage{subfigure}
%\usepackage{program}
%\usepackage{algorithmic}
\usepackage[boxed]{algorithm2e}
\usepackage{makecell}

@@

\subsection{Liveness Analysis for lazy languages}
\label{sec:liveness-analysis}
Figure~\ref{fig:live-judge}  describes  our  analysis  which  has  two
parts. The function $\mathit{ref}$,  takes  an application $s$ and
a demand $\sigma$, and returns  the incremental liveness generated for the
free  variables   of  $s$  due  to  the   application.   The  function
$\mathcal{L}$  uses   $\mathit{ref}$  to  propagate   liveness  across
expressions.

@@

@@

\subsection{Generating Liveness Equations}
The  program  in  Figure~\ref{fig:mot-example2}  serves as  a  running
example.   Given a  function $\mathit{f}$,  we now  describe  how to
generate    equations    for     the    demand    transformation
\Lfonly$_\mathit{f}$.   Starting with a  symbolic demand  $\sigma$, we
determine  \Lfun{e_{\mathit{f}}}{\sigma}{\Lfonly}.  In  particular, we
consider $\Lanv{1}{x_i}$,  the  liveness of    the
$i^{\mbox{\footnotesize  th}}$
parameter $x_i$ at the  program point at the  beginning of
$e_{\mathit{f}}$ (assumed  to be  $\pi_1$). By the rule {\sc
  live-define}, this should be the
same as $\Lf{f}{i}{\sigma}$. Applying this to  \length, we have:
\begin{eqnarray*}
&& \Lanv{1}{\xl} = \clazy\sigma \cup \acdr\Lf{\length}{1}{\clazy\sigma}
  \cup \clazy\Lf{\length}{1}{\clazy\sigma}
\end{eqnarray*}
and  the only equation defining \Lfone{\length} is:
  \begin{eqnarray*}
   && \Lf{\length}{1}{\sigma}
    =  \clazy\sigma \cup \acdr\Lf{\length}{1}{\clazy\sigma} \cup
\clazy\Lf{\length}{1}{\clazy\sigma}
\end{eqnarray*}

@@

In summary, the equations generated during liveness analysis are:
\begin{enumerate}
\item   For    each   function   $\mathit{f}$,    equations   defining
\Df{f}{i} for use by \Lfonly$_{\mathit f}$.
\item  For  each   function  $\mathit{f}$, an  equation  defining
the summary demand  $\sigma_{\mathit f}$ on $e_f$.
\item For each function $\mathit{f}$ (including $\mainpgm$ for
  $e_\mainpgm$)
an equation defining
  liveness at each GC point of $e_{\mathit f}$.
\end{enumerate}
\subsection{Solving liveness equations---the grammar
interpretation}\label{sec:grammar-formulation}
The  equations above can now be re-interpreted as a
context-free grammar (CFG)  on the alphabet $\lbrace\acar, \acdr,
\bcar, \bcdr, \clazy\rbrace$.  Let  \var{$X$} denote the non-terminal
for
a variable  $X$ occurring on  the LHS of the  equations generated
from the analysis.  We can  think of the resulting productions as
being associated with several  grammars, one for each non-terminal
\var{\Lanv{i}{x}} regarded as a start symbol.  As an example, the
grammar    for   \var{\Lanv{1}{\xl}}   comprises    the   following
productions:
 \begin{eqnarray*}
\var{\Lanv{1}{\xl}}  &\rightarrow&   \clazy\sigma_{\length} \mid \acdr
\var{\Df
  {\length}{1}}{\clazy\sigma_{\length}} \mid
  \clazy\var{\Df{\length}{1}}{\clazy\sigma_{\length}} \\
  \var{\Df{\length}{1}} &\rightarrow& \clazy \mid
  \acdr\var{\Df{\length}{1}}\clazy
       \mid \clazy\var{\Df{\length}{1}}{\clazy}\\
\langle {\sigma_{\length}} \rangle
&\rightarrow&
\epsilon  \mid \clazy\langle{\sigma_{\length}}\rangle
\end{eqnarray*}
Other equations can be converted similarly.  The language generated by
\var{\Lanv{i}{x}},  denoted $\mathscr{L}(\var{\Lanv{i}{x}}) $,  is the
desired  solution of  \Lanv{i}{x}.  However,  recall from  our earlier
discussion that the decision problem  that we are interested in during
garbage collection at a program point $\pi_i$ is:
\begin{quote}
Let  $x.\alpha$ be  a forward  access  path consisting  only of  edges
\acar\    and   \acdr\    (but    not   \bcar, \bcdr\ or \clazy).     Let
$\mathscr{L}(\var{\Lanv{i}{x}}) \stackrel{*}{\hookrightarrow} \sigma$,
where $\sigma$ consists  of forward paths only. Then  does $\alpha \in
\sigma$?
\end{quote}

@@

@@

\begin{figure}[t!]
\hspace*{1cm}{\epsfig{file=example-automata.eps, height=2.75cm}}
\caption{(a) The grammar rules for $\langle \Df{\length}{1'}\rangle$
converted into
  an automaton, and simplified in   (b). (c) and (d) show the automata
  for   \Lanv{9}{\pa}.}\label{fig:example-automata}
\figrule
\end{figure}

@@

\begin{figure*}[t!]
\label{fig:algos}
\hfill
\subfigure[]{\includegraphics[width=.30\linewidth]{lgc-scheme-1}}
\hfill
\subfigure[]{\includegraphics[width=.40\linewidth]{lgc-scheme-2}}
\hfill
\caption{(a) LGC with closure collection based on liveness.  (b)
  LGC with reachability-based closure collection.}

\end{figure*}

Figure~\ref{fig:algos}  shows   a  liveness-based  garbage  collection
scheme.   The algorithm  assumes that  checking for  liveness involves
traversing the  liveness-DFAs stored as  a table and does  not mention
this explicitly.  Starting  with the root set, each  live reference is
explored   using  $\mathit{  lgcCopy}$.    Copying  \CONS\   cells  is
simple---it  just involves  copying  the cell  itself and  recursively
copying the \CAR\ and  the \CDR\ fields.  Liveness-based collection of
closures  is more  complicated since  our analysis  only  computes the
liveness of evaluated objects and not closures.

@@

\subsection{Liveness-based collector with reachability-based closure
  collection}  Since the first  algorithm has  significant bookkeeping
overheads,  we have  a  second GC  scheme  that not  use liveness  for
copying closures.  In this  scheme, the liveness automata is consulted
only for \CONS\ cells; we assume everything under a closure to be live
and copy everything  that is reachable.

@@

\newcommand{\hgt}{3cm}
\begin{figure*}[t]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}c@{}}
\hskip -4mm{\epsfig{file=sudoku.eps, height=\hgt}}%, width=4.10cm}}
&
{\epsfig{file=lcss.eps, height=\hgt}}%, width=4.10cm}}
\\
\hskip -4mm{\epsfig{file=gc_bench.eps, height=\hgt}}%
 &
{\epsfig{file=nperm.eps, height=\hgt}}%
\\
\hskip -4mm{\epsfig{file=fibheap.eps, height=\hgt}}%
&
{\epsfig{file=fibheap.eps, height=\hgt}}%%%Add knightstour.eps
\\
\hskip -4mm{\epsfig{file=nqueens.eps, height=\hgt}}%
&
{\epsfig{file=treejoin.eps, height=\hgt}}%
\end{tabular}\vskip -10mm
 \caption{Memory usage  snapshot of  programs.  The  blue and  the red
   curves indicate the  number of cons cells in  the active semi-space
   for  LGC and  RGC  respectively.  The  black  curve represents  the
   number of reachable cells and  the grey curve represents the number
   of cells that are actually live  (of which liveness analysis does a
   static approximation).  x-axis is  the time  measured in  number of
   cons-cells allocated (scaled down by  factor $10^5$). y-axis is the
   number of cons-cells (scaled down by factor $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}
The  table  in  Figure~\ref{fig:experimental-results} shows  that  LGC
improves the  garbage collection time  and efficiency for most  of the
programs.  Even for  programs where the time taken is  more it is very
competitive.  Memory usage  graphs are shown for  select benchmarks in
Figure~\ref{fig:memory-usage}.  As  the number of  garbage collections
tend  to  be  very large  our  graphs  show  only  a window  which  is
representative of the behavior for  that particular benchmark.  In all
the programs we can see that  the curve corresponding to LGC regularly
dips below the RGC curve.

>>>>

\documentclass[9pt]{sigplanconf}
\usepackage{etex}
\usepackage{float}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage{boxedminipage}
\usepackage[T1]{fontenc}
\usepackage{epsfig}
\usepackage{multirow}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{schemepgm}
\usepackage{graphicx}
\usepackage{graphpap}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pstricks}
\usepackage{pst-text}
\usepackage{pst-node}
\usepackage{pst-tree}
\usepackage{pst-rel-points}
\usepackage{bcprules}
\usepackage{subfigure}
\usepackage[boxed]{algorithm2e}
\usepackage{makecell}

@@

\subsection{Liveness Analysis for lazy languages}
\label{sec:liveness-analysis}

Figure~\ref{fig:live-judge}  describes  our  analysis  which  has  two
parts. The function $\mathit{ref}$,  takes  an application $s$ and
a demand $\sigma$, and returns  the incremental liveness generated for the
free  variables   of  $s$  due  to  the   application.   The  function
$\mathcal{L}$  uses   $\mathit{ref}$  to  propagate   liveness  across
expressions.

@@

\subsection{Generating Liveness Equations}

@@

The  program  in  Figure~\ref{fig:mot-example2}  serves as  a  running
example.   Given a  function $\mathit{f}$,  we now  describe  how to
generate    equations    for     the    demand    transformation
\Lfonly$_\mathit{f}$.   Starting with a  symbolic demand  $\sigma$, we
determine  \Lfun{e_{\mathit{f}}}{\sigma}{\Lfonly}.  In  particular, we
consider $\Lanv{1}{x_i}$,  the  liveness of    the
$i^{\mbox{\footnotesize  th}}$
parameter $x_i$ at the  program point at the  beginning of
$e_{\mathit{f}}$ (assumed  to be  $\pi_1$). By the rule {\sc
  live-define}, this should be the
same as $\Lf{f}{i}{\sigma}$. Applying this to  \length, we have:
\begin{eqnarray*}
&& \Lanv{1}{\xl} = \clazy\sigma \cup \acdr\Lf{\length}{1}{\clazy\sigma}
  \cup \clazy\Lf{\length}{1}{\clazy\sigma}
\end{eqnarray*}
and  the only equation defining \Lfone{\length} is:
  \begin{eqnarray*}
   && \Lf{\length}{1}{\sigma}
    =  \clazy\sigma \cup \acdr\Lf{\length}{1}{\clazy\sigma} \cup
\clazy\Lf{\length}{1}{\clazy\sigma}
\end{eqnarray*}

@@

In summary, the equations generated during liveness analysis are:
\begin{enumerate}
\item   For    each   function   $\mathit{f}$,    equations   defining
\Df{f}{i} for use by \Lfonly$_{\mathit f}$.
\item  For  each   function  $\mathit{f}$, an  equation  defining
the summary demand  $\sigma_{\mathit f}$ on $e_f$.
\item For each function $\mathit{f}$ (including $\mainpgm$ for
  $e_\mainpgm$)
an equation defining
  liveness at each GC point of $e_{\mathit f}$.
\end{enumerate}
\subsection{Solving liveness equations---the grammar
interpretation}\label{sec:grammar-formulation}
The  equations above can now be re-interpreted as a
context-free grammar (CFG)  on the alphabet $\lbrace\acar, \acdr,
\bcar, \bcdr, \clazy\rbrace$.  Let  \var{$X$} denote the non-terminal
for
a variable  $X$ occurring on  the LHS of the  equations generated
from the analysis.  We can  think of the resulting productions as
being associated with several  grammars, one for each non-terminal
\var{\Lanv{i}{x}} regarded as a start symbol.  As an example, the
grammar    for   \var{\Lanv{1}{\xl}}   comprises    the   following
productions:
\begin{eqnarray*}
  \var{\Lanv{1}{\xl}}  &\rightarrow& 
  \clazy\sigma_{\length} \mid \acdr \var{\Df {\length}{1}}{\clazy\sigma_{\length}}  \\
  & & \mid
  \clazy\var{\Df{\length}{1}}{\clazy\sigma_{\length}} \\
  \var{\Df{\length}{1}} &\rightarrow& \clazy \mid
  \acdr\var{\Df{\length}{1}}\clazy
       \mid \clazy\var{\Df{\length}{1}}{\clazy}\\
\langle {\sigma_{\length}} \rangle
&\rightarrow&
\epsilon  \mid \clazy\langle{\sigma_{\length}}\rangle
\end{eqnarray*}
Other equations can be converted similarly.  The language generated by
\var{\Lanv{i}{x}},  denoted $\mathscr{L}(\var{\Lanv{i}{x}}) $,  is the
desired  solution of  \Lanv{i}{x}.  However,  recall from  our earlier
discussion that the decision problem  that we are interested in during
garbage collection at a program point $\pi_i$ is:
\begin{quote}
Let  $x.\alpha$ be  a forward  access  path consisting  only of  edges
\acar\    and   \acdr\    (but    not   \bcar, \bcdr\ or \clazy).     Let
$\mathscr{L}(\var{\Lanv{i}{x}}) \stackrel{*}{\hookrightarrow} \sigma$,
where $\sigma$ consists  of forward paths only. Then  does $\alpha \in
\sigma$?
\end{quote}

@@

\begin{figure}[t!]
  \begin{tabular}{cc}
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{15}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{7}{2}{\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-.5]{t1}{t1}
  \putnode{l1}{t1}{0}{10}{\acdr/\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-.5]{t2}{t2}
  \putnode{l2}{t2}{0}{10}{\clazy}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-.5]{t1}{t1}
  \putnode{l1}{t1}{0}{10}{\acdr}
\end{pspicture}
\\
(a) & (b) \\
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\Lanv{9}{\pa}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{12}{0}{\pscirclebox{\mbox{\ \ \ \ }}}
  \putnode{t3}{t2}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{5}{2}{\clazy}
  \ncline{t2}{t3}
  \putnode{l1}{t2}{5}{2}{\bcar}
  \nccurve[angleA=55, angleB=125, ncurv=3, nodesep=-.5]{t3}{t3}
  \putnode{l2}{t3}{0}{10}{\acdr}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\Lanv{9}{\pa}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t3}{t1}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t3}
  \putnode{l1}{t1}{5}{2}{\bcar}
  \nccurve[angleA=55, angleB=125, ncurv=3, nodesep=-.5]{t3}{t3}
  \putnode{l2}{t3}{0}{10}{\acdr}
\end{pspicture}
\\
(c) & (d)  \end{tabular}
\\ \\ \\ \\
\caption{(a) The  grammar rules for $\langle  \Df{\length}{1} \rangle$
  converted into an automaton, and simplified in (b). (c) and (d) show
  the   automata  for   \Lanv{9}{\pa}.   }\label{fig:example-automata}
\figrule
\end{figure}

@@

@@

  \SetStartEndCondition{ }{}{}%
  \SetKwProg{Fn}{procedure}{\string:}{}
  \SetKwFunction{Range}{range}%%
  \SetKw{KwTo}{in}\SetKwFor{For}{for}{\string:}{}%
  \SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
  \SetKwFor{While}{while}{:}{fintq}%
  \AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%
  \SetKwFunction{Lgc}{$lgc$}%
  \SetKwFunction{Copy}{$lgcCopy$}%
  \SetKwFunction{LCopy}{$lgcCopy$}%
  \SetKwFunction{RCopy}{$rgcCopy$}%

\begin{algorithm}[t]
  \Fn{\Lgc{}}
     {
       \For {each reference $\mathit{ref}$ in root set}
            {\Copy($\mathit{ref}$)\;}
            $\mathit{copyReferencesOnPrintStack()}$\;  
     }

     \Fn{\Copy{$\mathit{ref}$}}
        {
          \If{$\mathit{ref}$ is live}
             {
               $\mathit{newRef} = $  $\mathit{dupHeapCell(ref})$\;
               \If{cellType($\mathit{ref}$) is $cons$}
                  {
                    {
                      $\mathit{newCar}  = $  \Copy($\mathit{ref\!.car})$\;
                      $\mathit{newRef.car}   =  newCar$\;
                    }
                    {
                      $\mathit{newCdr}$ =   \Copy($\mathit{ref.cdr})$\;
                      $\mathit{newRef.cdr}   = \mathit{newCdr}$\;
                    }    
                  }
               \If {cellType($\mathit{ref}$) is closure}
                   {
                     $\mathit{newRef.arg1}$ =  \Copy$\mathit{(cell.arg1)}$\;
                     $\mathit{newRef.arg2}$ =  \Copy$\mathit{(cell.arg2)}$\;
                   }
             }
        }

        \caption{LGC with closure collection based on liveness.\label{algo:lgc-a}}
\end{algorithm}
Algorithm~\ref{algo:lgc-a}  describes  a  liveness-based  garbage  collection
scheme.   The algorithm  assumes that  checking for  liveness involves
traversing the  liveness-DFAs stored as  a table and does  not mention
this explicitly.  Starting  with the root set, each  live reference is
explored   using  $\mathit{  lgcCopy}$.    Copying  \CONS\   cells  is
simple---it  just involves  copying  the cell  itself and  recursively
copying the \CAR\ and  the \CDR\ fields.  Liveness-based collection of
closures  is more  complicated since  our analysis  only  computes the
liveness of evaluated objects and not closures.

@@

\subsection{Liveness-based collector with reachability-based closure
  collection}
\begin{algorithm}[t]
  \Fn{\Lgc{}}
     {
       \For {each reference $\mathit{ref}$ in root set}
            {copy($\mathit{ref}$)\;}
            $\mathit{copyReferencesOnPrintStack()}$\;  
     }
     \Fn{\LCopy{$\mathit{ref}$}}
        {
          \eIf{cellType$(\mathit{ref})$ is $cons$}
              {
                \eIf {$\mathit{ref}$ is live and has not been copied using
                  reachability}
                     {
                       $\mathit{newRef}$ =  $\mathit{dupHeapcell(ref})$\; 
                       $\mathit{ newCar}  =$  \LCopy$\mathit{(ref.car)}$\;
                       $\mathit{ newRef.car}   = \mathit{ newCar}$\;
                       $\mathit{ newCar}  =$  \LCopy$\mathit{(ref.cdr)}$\;
                       $\mathit{ newRef.cdr}   = \mathit{newCar}$\;
                     }
                     { $\mathit{makenull(ref)}$\;}}
              {
                \If {cellType$\mathit{(ref)}$ is closure and 
                  has not been copied}
                    { 
                      $\mathit{newRef}$  = \RCopy$(\mathit{ref})$\;
                    }
              }
        }
        \caption{LGC with reachability-based closure collection.\label{algo:lgc-b}}
\end{algorithm}
  Since the first  algorithm has  significant bookkeeping
overheads,  we have  a  second GC  scheme  that not  use liveness  for
copying closures.  In this  scheme, the liveness automata is consulted
only for \CONS\ cells; we assume everything under a closure to be live
and copy everything  that is reachable. Algorithm~\ref{algo:lgc-b}  describes  this scheme.

@@

\newcommand{\hgt}{3cm}
\begin{figure*}[t]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}c@{}}
\hskip -4mm{\epsfig{file=sudoku.eps, height=\hgt}}%, width=4.10cm}}
&
{\epsfig{file=lcss.eps, height=\hgt}}%, width=4.10cm}}
\\
\hskip -4mm{\epsfig{file=gc_bench.eps, height=\hgt}}%
 &
{\epsfig{file=nperm.eps, height=\hgt}}%
\\
\hskip -4mm{\epsfig{file=fibheap.eps, height=\hgt}}%
&
{\epsfig{file=sudoku.eps, height=\hgt}}%%%Add knightstour.eps
\\
\hskip -4mm{\epsfig{file=nqueens.eps, height=\hgt}}%
&
{\epsfig{file=treejoin.eps, height=\hgt}}%
\end{tabular}\vskip -10mm
 \caption{Memory usage  snapshot of  programs.  The  blue and  the red
   curves indicate the  number of cons cells in  the active semi-space
   for  LGC and  RGC  respectively.  The  black  curve represents  the
   number of reachable cells and  the grey curve represents the number
   of cells that are actually live  (of which liveness analysis does a
   static approximation).  x-axis is  the time  measured in  number of
   cons-cells allocated (scaled down by  factor $10^5$). y-axis is the
   number of cons-cells (scaled down by factor $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}
The  table  in  Figure~\ref{fig:experimental-results} shows  that  LGC
improves the  garbage collection time  and efficiency for most  of the
programs.  Even for  programs where the time taken is  more it is very
competitive.  Memory usage  graphs are shown for  select benchmarks in
Figure~\ref{fig:memory-usage}.  As  the number of  garbage collections
tend  to  be  very large  our  graphs  show  only  a window  which  is
representative of the behavior for  that particular benchmark.  In all
the programs we can see that  the curve corresponding to LGC regularly
dips below the RGC curve.
